### 2016-07-27 syntax.tex
updating syntax of data structures, and adding set-multiplicity data structures

\section{Secure Multiset Representation}
\def\pub{{\sf pub}}
\def\aux{{\sf aux}}

@@

\heading{Static (Multi)set-representations. }  We being by defining a general primitive that produces a representation of a given multiset, and provides a mechanism to answer a specified collection of questions about that multiset.  Informally, the representation should serve as a good approximation of the multiset, at least with respect to the allowed questions.
\begin{definition} 
Fix a universe~$\univ$ and an alphabet~$\Sigma$.  A \emph{static (multi)set-representation primitive} (or just set-representation) is a pair of algorithms $\setprim=(\Rep, \Qry)$
parameterized by a collection $\mathcal{S}$ of multisets, each consisting of elements from~$U$ and a collection
of supported queries $\mathcal{Q}=\{q \colon \mathcal{S} \to \mathbb{R}\}$, where:
\begin{itemize}
\item $\Rep$ is a randomized algorithm taking as input $S \in \mathcal{S}$,
and outputting a pair of strings $(M,\aux)$, the \emph{representation} $M \in \Sigma^*$ (of~$S$) 
and the \emph{auxilliary information} $\aux \in \bits^*$ (of~$M$).
\item $\Qry$ is a deterministic algorithm that takes as input $M, \aux$,
and a query $q\in \mathcal{Q}$, and returns value~$a$.
\end{itemize}
A set-representation is \emph{$\epsilon$-correct} if, for all $S \in \mathcal{S}$ and
all $q \in \mathcal{Q}$, it holds that
\begin{eqnarray}
\Pr[(M,\aux) \getsr \Rep(S) : \Qry(M, \aux, q) \neq q(S)] \leq \epsilon, \label{eqn:correctness}
\end{eqnarray}
%where $q(S)$ denotes the correct answer to the query $q$ on the set~$S$.
\end{definition}

@@

All of the constructions of set-representations that we will consider have the property that~$|M|$ does not depend on the coins of~$\Rep$. Thus we assume that~$|M|$ depends only on~$|S|$.  In general this need not be true, but the assumption does not restrict our syntax from capturing natural constructions (e.g., those seen in practice), and it makes it easier to state our security notions.
%
%\todo{Readdress issue of representation size.  A general statement might be something like `` The \emph{size} of a set-representation is the maximum (bit)length of any~$M$ that can be output by $\Rep$.''  But all of the natural constructions will have representation whose size will not depend on the coins of $\Rep$, i.e. they depend on $|S|$, number of hash functions, desired error rate. }

@@

\heading{Examples.}
It is not hard to see Bloom filters are one possible instantiation a set-representation primitive.
Fix an $n \geq 0$, let $\mathcal{S}=[\univ]^n$, and fix $\Sigma=\bits$.  
Next, let $\calQ$ be a set of predicates $q_x$, one for each $x \in \univ$,
such that $q_x(S)=1$ iff $x \in S$.  Define algorithm~$\Rep$, with functions $k=k(|S|), m=m(|S|)$ and a description of a hash function family~$\mathcal{H}=\{h \colon \univ \to \{1,2,\ldots,m\}\}$ hardwired into it, to sample hash functions $h_1, \ldots, h_k$ from~$\mathcal{H}$ and then computes an $m$-bit array in the standard way. It sets $\aux = (h_1, \ldots, h_k)$ and $M$ to be the bit array, and returns these.  The algorithm~$\Qry$ is defined so that, on input $(M,\aux,q_x)$, it returns 1 iff $M[h_1(x)]=1 \wedge M[h_2(x)]=1 \wedge \cdots \wedge M[h_k(x)]=1$ holds.
  %\tsnote{If $\mathcal{Q}=U$ and we use the
%  semantics you suggest, then isn't Equation~\ref{eqn:correctness}
%  both correctness and soundness?  If so, I don't see how this
%  captures the classical BF, which requires $\epsilon=0$ for one kind
%  of error, but allows $\epsilon>0$ for the other. } 
%\jnote{The definition does not distinguish
%  between elements in the set (for which $q(S)=1$) or outside the set (for which
%  $q(S)=0$). But a Bloom filter does satisfy Equation~\ref{eqn:correctness} in either case.}
%  \jnote{In that sense, I am using ``soundness'' and ``correctness'' interchangeably here.}
%  \jnote{I'm not tied to allowing false negatives when we specialize the definition to the case
%  of set-membership queries. But in the abstract definition it seems unclear why one would choose
%  some queries on which error is ok and other queries where error is not allowed. Though if one
%  wanted to be general, one could simply have two (disjoint) sets of queries $\mathcal{Q}_1, \mathcal{Q}_2$
%  where error is allowed on the former but not on the latter.}
We note that while our correctness condition does not distinguish among types of errors, it also does not preclude instantiations (like the classical Bloom filter) that admit only one type of error, e.g. false-positives.

>>>>

\section{Secure Data Structures}
\def\pub{{\sf pub}}
\def\aux{{\sf aux}}

@@

\heading{(Set-multiplicity) Data Structures.}  We being by defining a general primitive that produces a representation of a given multiset, and provides a mechanism to answer a specified collection of questions about that multiset.  Informally, the representation should serve as a good approximation of the multiset, at least with respect to the allowed questions.
\begin{definition} 
Fix a non-empty set~$\univ$. A \emph{static data structure} (over universe~$\univ$)
is a tuple $\setprim=(\Sigma,\Gamma,\calS,\calQ,\Rep, \Qry)$ where:
\begin{itemize}
\item $\Sigma$ is the \emph{representation alphabet}
\item $\Gamma$ is the set of possible \emph{query answers}
\item $\calS$ is a set  of multisets over universe~$\univ$
\item $\mathcal{Q}=\{q \colon \mathcal{S} \to \Gamma\}$ is a set of supported \emph{queries}
\item $\Rep \colon \mathcal{S} \to \Sigma^* \times \bits^*$ is a randomized \emph{representation algorithm}, taking as input $S \in \mathcal{S}$, and outputting a \emph{representation} $M \in \Sigma^*$ and some \emph{auxilliary information} $\aux \in \bits^*$
\item $\Qry\colon \Sigma^* \times \bits^* \times \calQ \to \Gamma$ is a deterministic \emph{query-evaluation} algorithm, taking as input $M, \aux, q$, and returns \emph{answer} $a \in \Gamma$.
\end{itemize}
Let $\mathrm{Err}(S,q,M,\aux)$ be a predicate that is true iff $\Qry(M, \aux, q) \neq q(S)$.  Then a data structure
\emph{$\epsilon$-correct}, or has (worst-case) error probability $\epsilon$, if
\[
\max_{S,q}\Pr[(M,\aux) \getsr \Rep(S) : \mathrm{Err}(S,q,M,\aux)=1 ] \leq \epsilon, \label{eqn:correctness}
\]
where the maximum is over $S \in \mathcal{S}$ and $q \in \mathcal{Q}$.  Relatedly, it has \emph{error rate} $\tilde{\epsilon}$ if
\[
\max_{S}\Pr[(M,\aux) \getsr \Rep(S);\, q \getsr \calQ : \mathrm{Err}(S,q,M,\aux)=1] \leq \tilde{\epsilon} \label{eqn:error_rate}
\]
\end{definition}

@@

All of the data structures that we will consider have the property that~$|M|$ does not depend on the coins of~$\Rep$. Thus we assume that~$|M|$ depends only on~$|S|$.  In general this need not be true, but the assumption does not restrict our syntax from capturing natural constructions (e.g., those seen in practice), and it makes it easier to state our security notions.

Our particular focus will be on data structures that support multiplicity queries, of which set-membership queries are a special case.  
\begin{definition} A \emph{set-multipicity data structure} is a data structure with $D=\mathbb{N}$ and where $\calQ=\{q_x \,|\, x \in \univ\}$ where $q_x(S)$ is defined to be the multiplicity of~$x$ in the multiset~$S$.
\end{definition}
%
%\todo{Readdress issue of representation size.  A general statement might be something like `` The \emph{size} of a set-representation is the maximum (bit)length of any~$M$ that can be output by $\Rep$.''  But all of the natural constructions will have representation whose size will not depend on the coins of $\Rep$, i.e. they depend on $|S|$, number of hash functions, desired error rate. }

@@

\heading{Examples.}
It is not hard to see Bloom filters are one possible instantiation a set-multiplicity data structure.
Fix an $n \geq 0$, let $\mathcal{S}=[\univ]^n$, and let $\Sigma=\Gamma\bits$.  
Let $\calQ$ be a set of predicates $q_x$, one for each $x \in \univ$,
such that $q_x(S)=1$ iff $x \in S$.  Define algorithm~$\Rep$, with functions $k=k(|S|), m=m(|S|)$ and a description of a hash function family~$\mathcal{H}=\{h \colon \univ \to \{1,2,\ldots,m\}\}$ hardwired into it, to sample hash functions $h_1, \ldots, h_k$ from~$\mathcal{H}$ and then computes an $m$-bit array in the standard way. It sets $\aux = (h_1, \ldots, h_k)$ and $M$ to be the bit array, and returns these.  The algorithm~$\Qry$ is defined so that, on input $(M,\aux,q_x)$, it returns 1 iff $M[h_1(x)]=1 \wedge M[h_2(x)]=1 \wedge \cdots \wedge M[h_k(x)]=1$ holds.
  %\tsnote{If $\mathcal{Q}=U$ and we use the
%  semantics you suggest, then isn't Equation~\ref{eqn:correctness}
%  both correctness and soundness?  If so, I don't see how this
%  captures the classical BF, which requires $\epsilon=0$ for one kind
%  of error, but allows $\epsilon>0$ for the other. } 
%\jnote{The definition does not distinguish
%  between elements in the set (for which $q(S)=1$) or outside the set (for which
%  $q(S)=0$). But a Bloom filter does satisfy Equation~\ref{eqn:correctness} in either case.}
%  \jnote{In that sense, I am using ``soundness'' and ``correctness'' interchangeably here.}
%  \jnote{I'm not tied to allowing false negatives when we specialize the definition to the case
%  of set-membership queries. But in the abstract definition it seems unclear why one would choose
%  some queries on which error is ok and other queries where error is not allowed. Though if one
%  wanted to be general, one could simply have two (disjoint) sets of queries $\mathcal{Q}_1, \mathcal{Q}_2$
%  where error is allowed on the former but not on the latter.}
We note that while our correctness condition does not distinguish among types of errors, it also does not preclude instantiations (like the classical Bloom filter) that admit only one type of error, e.g. false-positives.
